---
title: "MachineLearning.Rmd"
output: html_document
---
Loading data. 
```{r cache = TRUE}
fileURL<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileURL, destfile = "pml-training.csv", method = "curl")
training <- read.csv("pml-training.csv", header = TRUE,  na.strings= c("NA", "#DIV/0!", ""))

fileURL1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileURL1, destfile = "pml-testing.csv", method = "curl")
testing <- read.csv("pml-testing.csv", header = TRUE, na.strings= c("NA", "#DIV/0!", ""))
```

Libraries and seed for reproducibility. Paritioning the training dataset into two subsets, one for training and one for testing. 
```{r}
library(caret)
library(randomForest)
library(rpart)
library(rattle)
set.seed(1234)
inTrain = createDataPartition(y = training$classe, p = 0.6, list = F)
myTrain = training[inTrain, ]
myTest = training[-inTrain, ]
dim(testing)
```



Remove all NA's above a certain threshold, 60 percent. 
```{r }
myTrain = myTrain[c(-1)]
trainingV3 <- myTrain
for(i in 1:length(myTrain)) { 
        if( sum( is.na( myTrain[, i] ) ) /nrow(myTrain) >= .6 ) { #if n?? NAs > 60% of total observations
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTrain[i]), names(trainingV3)[j]) ) ==1)  { 
                trainingV3 <- trainingV3[ , -j] #Remove that column
            }   
        } 
    }
}
myTrain <- trainingV3
clean1 <- colnames(myTrain)
clean2 <- clean1[1:58]

myTest <- myTest[clean1]
testing <- testing[clean2]



#Coercion
for (i in 1:length(testing) ) {
        for(j in 1:length(myTrain)) {
        if( length( grep(names(myTrain[i]), names(testing)[j]) ) ==1)  {
            class(testing[j]) <- class(myTrain[i])
        }      
    }      
}
testing <- rbind(myTrain[2, -59] , testing) 
testing <- testing[-1,]

```

## Machine Learning
Random Forest model is used to test the data.
```{r}
model <- randomForest(classe ~ . - classe,data = myTrain)

```

## Predictions
Used testing training data set to see effectiveness
```{r}
prediction <- predict(model, myTest, type = "class")
confusionMatrix(prediction, myTest$classe)

```

## Final Predictions
Predicted using untouched, testing dataset.
```{r}
testpredictions <- predict(model, testing, type = "class")
confusionMatrix(prediction, myTest$classe)
```
